---
title: "COmparison of Trial Measurement Methodologies"
output:
  html_document:
    df_print: paged
---

The objective of this analysis is to compare some standard trial measurement approaches for a specific scenario and identify some pros and cons.

*Problem Overview
A retail chain uses a recommender system to suggest products that the customer might not have purchased earlier. (This notebook won't delve into the specifics of how such a system might work). THe hypothesies is that these customers who receive recommendations spend more than the other customers. THis motebook aims to design an approach that might identify such an uplift.


```{r, echo = F}
library(data.table)
library(stringi)
library(optiRum)

```


# Approach

```{r}
# We start by settinf up the trial config

config <- list(
  ttl_customers = 10000,
  trial_smp_size = 1000,
  prop_ctrl_grp = 0.5,
  num_trials = 1,
  sales_param_rng = c(10,150),
  sd_sales_param = 5,
  trial_duration_wks = 10,
  pre_trial_weeks = 5
)

```


Defining  set of total customers with a 'sales parameter' for each cutomer. This parameter would define how much they are likely to spend each week at this store.

```{r}

# Preparing Trial Data for the analysis

cust_dt <- data.table(cust_id = paste0('CN',1:config$ttl_customers),
                       sales_param = runif(min = config$sales_param_rng[1],
                                           max = config$sales_param_rng[2],n = config$ttl_customers))

# Create subset of customers for trial
trial_cust <- sample(cust_dt[,unique(cust_id)], config$trial_smp_size, replace = F)
trial_cust_dt <- cust_dt[cust_id %in% trial_cust,]
# Splitting between treatment and control groups
trial_cust_dt[,treatment := ifelse(cust_id %in% sample(unique(trial_cust_dt$cust_id),
                                                       size = config$prop_ctrl_grp*nrow(trial_cust_dt),replace = F),0,1)]

ttl_dt <- CJ.dt(trial_cust_dt,data.table(week = c(1:config$trial_duration_wks)))
ttl_dt[,sales := rnorm(sales_param,mean = sales_param, sd = config$sd_sales_param)]
ttl_dt[,post := ifelse(week > config$pre_trial_weeks,1,0)]

head(ttl_dt)

```

# Testing Methodologies 
Now that we have a sample trial dataset ready, we will set up functions to test uplifts. For this analysis, we will use two approaches - the ratio method and the Difference in Differences (DiD) method.

Starting with defining the ratio method - 

```{r}

calc_ratio_up <- function(config,dat){
  sum_dt <- dat[,.(mean_val = mean(sales)),.(post,treatment)]
  uplift <- sum_dt[post == 1 & treatment == 1, mean_val] - 
            sum_dt[post == 0 & treatment == 1, mean_val]/sum_dt[post == 0 & treatment == 0, mean_val]*sum_dt[post == 1 &    treatment == 0, mean_val]
  
  return(uplift)
}

```

And now for the simple DiD implementation

```{r}

calc_did_up <- function(config, dat){
  
  fit1 <- lm(formula = 'sales ~ treatment + post + treatment:post', data = dat)
  uplift <- summary(fit1)$coefficients[4]
  conf_int <- confint(fit1)[c(4,8)]
  
  return(list(uplift = uplift,
              rng = conf_int))
}

```

# A/A Testing

Before start measuring some actual uplifts, we need to test a scenario with regular data to make sure that we are not unintentionally capturing any stray signals.
To start with we'll test both our methodologies with the current dataset (which shouldnt result in any uplift),

```{r}

uplift_rat <- calc_ratio_up(config, ttl_dt)
uplift_rat

```

Now with the DiD

```{r}

uplift_rat2 <- calc_did_up(config, ttl_dt)
uplift_rat2$uplift
uplift_rat2$rng

```

Both methods return a negligible effect and the confidence interval for the did measurement includes 0, so the results are what we would expect to see.

# Incorporating an additive uplift

Lets assume that customers subjected to the product recommendations spend about $10 more than if they were not sent those recommendations.Also, we'll add in some noise to the data to see how difficult it is to retrieve this information.
Adding some uplift parameters to the config.

```{r}
config[['uplift_mean']] = 10
config[['uplift_sd']] = 5
```

Now updating the trial data with the uplift info

```{r}
act_uplift <- rnorm(ttl_dt[post ==1 & treatment ==1,.N],config$uplift_mean, config$uplift_sd)
ttl_dt[, uplift := 0]
ttl_dt[post == 1 & treatment == 1, uplift := act_uplift]
```

```{r}
#Calculating updated sales
ttl_dt[,sales := rnorm(sales_param,mean = (sales_param + uplift), sd = config$sd_sales_param)]


uplift_new <- calc_ratio_up(config,ttl_dt)

uplift_new2 <- calc_did_up(config,ttl_dt)
uplift_new2$uplift
uplift_new2$rng

```

